# Calculation

This tutorial will help you to calculate the lagged correlation between two time series. It will use `calculate_lagged_correlation` function of `lagci` package.

## Prepare your data

You can get started with the example data in `lagci` package. And the format of `time` column is `POSIXct`

```{r,eval=TRUE}
library(lagci)

data("heart_data")
data("step_data")

head(heart_data)
head(step_data)

class(heart_data$time)
class(step_data$time)
```

## Calculation

Calculate with `calculate_lagged_correlation` function of `lagci` package. \### x, y

Type: Numeric vectors

Purpose: The two time-series to be compared.

Details: Each vector represents the observed values of one variable. Internally, both vectors are standardized (centered and scaled to unit variance) to make the correlation unbiased by different magnitudes or units.

### time1, time2

Type: POSIXct vectors (or convertible to POSIXct)

Purpose: Timestamps corresponding to x and y.

Role: Used to align the two time-series onto a shared, regularly spaced timeline before computing correlations.

### time_tol

Type: Numeric (hours)

Default: 1

Meaning: The maximum time lag to evaluate in both positive and negative directions.

Unit: Hours. For example:

time_tol = 1 → ±1 hour

time_tol = 0.5 → ±30 minutes

time_tol = 1/60 → ±1 minute

### step

Type: Numeric (hours)

Default: 1/60 (i.e., 1 minute)

Meaning: The resolution of the regular time grid used for alignment.

Unit: Hours. For example:

step = 1 → 1-hour sampling

step = 1/60 → 1-minute sampling

step = 1/3600 → 1-second sampling

### min_matched_sample

Type: Integer

Default: 10

Purpose: Minimum number of overlapping (non-missing) samples required after alignment. If fewer matched samples remain, the function returns NULL to avoid unstable correlations.

### progressbar

Type: Logical

Default: TRUE

Purpose: Whether to display progress information. Particularly useful when applying the function to many pairs of series in batch analyses.

### align_method

Type: Character string

Options: "linear" or "constant"

Default: "linear"

Purpose: Method used for interpolation when mapping the irregular time-series onto the regular grid.

"linear": Interpolates linearly between neighboring observations.

"constant": Uses the nearest neighbor (step-wise constant).

### cor_method

Type: Character string

Options: "spearman" or "pearson"

Default: "spearman"

Purpose: The statistical correlation measure.

"pearson": Measures linear correlation between the raw values.

"spearman": Measures rank correlation, robust to monotonic but non-linear relationships and less sensitive to outliers.

### Calculation

```{r,eval=TRUE}
result_1 <- lagci::calculate_lagged_correlation(
  x = heart_data$heart,
  y = step_data$step,
  time1 = heart_data$time,
  time2 = step_data$time,
  time_tol = 0.5,
  step = 1/60,
  align_method = "linear",
  cor_method = "spearman"
)

print(result_1)
```

```{r,eval=TRUE}
result_2 <- lagci::calculate_lagged_correlation(
  y = heart_data$heart,
  x = step_data$step,
  time2 = heart_data$time,
  time1 = step_data$time,
  time_tol = 0.5, # 
  step = 1/60,
  align_method = "linear",
  cor_method = "spearman"
)

print(result_2)
```
