# Installation of lagcishiny

You can install `lagcishiny` successfully follow the following steps:

## Installation

```{r,eval=FALSE}
if (!require("pak", quietly = TRUE)) 
  install.packages("pak")
pak::pak("jaspershen-lab/lagci")
pak::pak("jaspershen-lab/lagcishiny")
```

### Install TinyTex

If you want to export reports in PDF format, you should install `TinyTex`. `TinyTex` is a simple `Tex` suite to compile `PDF` files, it is convenient to use in PC.

There are two ways to install `TinyTex` for different users respectively.

#### Directly install TinyTex

After installing `lagcishiny`, you can install `TinyTex` via this command:
```{r,eval=FALSE}
tinytex::install_tinytex()
```

Check if the `TinyTex` is installed successfully:
```{r,eval=FALSE}
tinytex::is_tinytex()
tinytex::tlmgr_version()
```

#### Locally install TinyTex

You can download the correct `TinyTex` release at [Github release](https://github.com/rstudio/tinytex-releases), then install it via:
```{r,eval=FALSE}
tinytex:::install_prebuilt(pkg = "PATH/TO/YOUR/FILE")
```

Check if the `TinyTex` is installed successfully:
```{r,eval=FALSE}
tinytex::is_tinytex()
tinytex::tlmgr_version()
```

## Session information

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```
